ms2

### 2017-06-20 ms2/ms2.Rmd
merging Staph, filtering n

```{r load_abg, cache.extra=md5sum('bug_drug.tsv')}
resistance_groups = read_tsv('resistance_groups.tsv')
bug_names = read_tsv('bug_names.tsv')

raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by=c('zipcode', 'state')) %>%
  right_join(resistance_groups, by='drug') %>%
  select(bug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

possible_bug_drug = read_tsv('bug_drug.tsv')

# Keep only the abundant bug/drug combinations
#min_n_states = 25
#good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
#  select(bug, drug_group, state) %>%
#  distinct %>%
#  count(bug, drug_group) %>% ungroup() %>% 
#  rename(n_places=n) %>%
#  filter(n_places >= min_n_states)

abg = raw_abg %>%
  #right_join(good_combos, by=c('bug', 'drug_group')) %>%
  right_join(possible_bug_drug, by=c('bug', 'drug_group')) %>%
  rename(bug_long=bug) %>% left_join(bug_names, by='bug_long') %>% select(-bug_long) %>%
  filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))

# check that all the short names made it in
if (any(is.na(abg$bug))) stop('missing bug names')

summarize_abg = function(df, place_name) {
    group_by_(df, 'bug', 'drug_group', place_name) %>%
    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
              n_place_antibiograms=n()) %>%
    mutate(place_weight=sqrt(n_place_antibiograms))
}

@@

I started by considering these bug/drug combinations on the basis of their clinical relevance and plausibility:

@@

>>>>

```{r load_abg, cache.extra=md5sum('bug_drug.tsv')}
resistance_groups = read_tsv('resistance_groups.tsv')
bug_names = read_tsv('bug_names.tsv')
bug_groups = read_tsv('bug_groups.tsv')
possible_bug_drug = read_tsv('bug_drug.tsv')

raw_abg = read_tsv('../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by=c('zipcode', 'state')) %>%
  right_join(resistance_groups, by='drug') %>%
  select(bug, drug, drug_group, state, hrr, percent_nonsusceptible, n_isolates)

# Keep only the abundant bug/drug combinations
min_n_states = 35
good_combos = semi_join(raw_abg, possible_bug_drug, by=c('bug', 'drug_group')) %>%
  select(bug, drug_group, state) %>%
  distinct %>%
  count(bug, drug_group) %>% ungroup() %>% 
  rename(n_places=n) %>%
  filter(n_places >= min_n_states)

abg = raw_abg %>%
  #mutate(raw_bug=bug) %>%
  # shorten the bug names
  left_join(bug_names, by='bug') %>%
  select(-bug) %>% rename(bug=bug_short) %>%
  # group the different S. aureus into meta-groups depending on the drug
  left_join(bug_groups, by=c('bug', 'drug_group')) %>%
  mutate(bug_group=if_else(is.na(bug_group), bug, bug_group)) %>%
  select(-bug) %>% rename(bug=bug_group) %>%
  right_join(good_combos, by=c('bug', 'drug_group')) %>%
  # keep only particular bug/drug combos
  right_join(possible_bug_drug, by=c('bug', 'drug_group')) #%>%
  #filter(!is.na(percent_nonsusceptible), !is.na(n_isolates))

# get the median number of isolates for each bug/drug
median_n_isolates = abg %>%
  filter(!is.na(n_isolates)) %>%
  group_by(bug, drug_group) %>%
  summarize(median_n_isolates=as.integer(median(n_isolates))) %>%
  ungroup()

# replace the NA number of isolates with the median for that bug/drug
abg %<>%
  left_join(median_n_isolates) %>%
  mutate(n_isolates=if_else(is.na(n_isolates), median_n_isolates, n_isolates)) %>%
  select(-median_n_isolates)

# check that all the short names made it in
for (x in c('bug', 'n_isolates', 'percent_nonsusceptible')) {
  if (any(is.na(abg[[x]]))) stop(sprintf('NAs in field "%s"', x))
}

summarize_abg = function(df, place_name) {
    group_by_(df, 'bug', 'drug_group', place_name) %>%
    summarize(mean_percent_nonsusceptible=weighted.mean(percent_nonsusceptible, sqrt(n_isolates)),
              n_place_antibiograms=n()) %>%
    ungroup() %>%
    mutate(place_weight=sqrt(n_place_antibiograms))
}

@@

I started by considering these `r nrow(possible_bug_drug)` bug/drug combinations on the basis of their clinical relevance and plausibility:

@@

```{r}
x = abg %>% 
  select(bug, drug_group, state) %>%
  distinct() %>%
  count(bug, drug_group) %>% ungroup() %>% rename(n_places=n)
results %>% filter(model=='spearman') %>%
  left_join(x, by=c('bug', 'drug_group')) %>%
  ggplot(aes(x=n_places, y=estimate)) +
  geom_point()
```
